---
title: "WUP_04"
author: "Matt Diaz"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS = read.csv("GSScsv.csv")
```

# Introduction

```{r}
data <- read.csv(file="GSScsv.csv")
```

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We want to predict the WORDSUM based on DEGREE.

WORDSUM is a numerical variable.
DEGREE is a categorical variable.

So we use the numerical~cat helper file. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
GSS = read.csv("GSScsv.csv")
boxplot(WORDSUM~DEGREE, data=GSS)
```

According to this boxplot, there are clear differences with every median value between every degree within LT High School. The lowest degree has a value of 8 and the median value contains a value of 11. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=data)
```

There is a difference in the mean based on the degree with the lowest being the LT high school with an average score of 7.75 which will be compared to the Graduate which contains an average score of 10.09.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```

Here we can identify that that p-value is 0.000236 which is less than 0.05 so we will reject the null hypothesis. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The question that was originally stated was asking us if there is a difference between average score on the WORDSUM test and the level of which this test was scored. When we take a look at the boxplot we can see that the median values are all different as do the minimum and maximum values. Based on all of this information that we have received from the plots and data, we can reject the null hypothesis and we see that there is a significant between degree and WORDSUM test score. 
